---
title: "Insight Matching Algo"
author: "Juan Carpio"
date: "12 December 2017"
output: html_document
---


# match the datasets to play with the matrix pilot..


This code matches any data set containing addresses with the LLPG.

In order to run the code, some of the following part of the code mat need to be edited:
  
  + Define working directory.
  + Upload raw data.
  + Define variable labels.
  + Define "Working address" and relevand regords.
  + List variables from matching dataset that need to be retained.


# CODE TO BE EDITED (if needed). 

### 1- Define working directory

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE, eval=TRUE,echo=FALSE}
rm(list=ls())
#opts_chunk$set(root.dir = "~/Z:/Insight Hub/LLPG/Matching_LLPG")
setwd("Z:/Insight Hub/LLPG/Matching_LLPG")

```

### 2- Uploading raw data

The dataset to be matched with the LLPG needs to be stored in a dataframe called "b"
```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}
library("data.table")
# Upload data sets 
a<- fread("//Bbd-dfs-p-fil05/cscom/Policy and Research/Research, Intelligence and Customer Insight/Insight Hub/LLPG/UNI72LIVE20171019.csv", na.strings=c("")) 

## MODIFY ##
b<- fread("//dscs/cscom/Policy/Corp Information/Olypmic Hosts NKM (secure)/2016_Resident_Matrix/RawData/GPRegister/GP_RAW_March2016.csv", na.strings=c(""), 
          colClasses=c(NHS_NUMBER="character"))# GP Registry
names(b)

```

### 3- Define variable labels

The variables from the output files will identify the original data set that they were created from by adding a suffix at the end of the name of the variable. For instance, we can define "zt" as the suffix identifying variables from the Gazetteer data set (LLPG), so the variable address will be called "adr_gz". The labels for the suffix in each data set can be defined below.  

```{r , include=TRUE, warning=FALSE, cache=TRUE, message=FALSE,echo=FALSE}

label.a<- "gz" # llpg
  
label.b<- "gp" # GP
  
```


### 4- Defining "Working address"

The matching process is mainly based in a variable called "work.adr" (working address) that needs to be created at the beginning of the process. This variable contains the property/premise number, and the street name (e.g. 104 Main Street). It may require joining different columns depending on the data set.    

#### 4.1- Dataset (b) to be match with the LLPG (**ADAPT THIS SECTION**)

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}
library("dplyr")
b.working<- b %>% select(everything(),id=NHS_NUMBER, name=FORENAME, surname=SURNAME,
                         for.surname=PREVIOUS_SURNAME,gen=SEX,
                         dob=DATE_OF_BIRTH,adr1=PREMISES,
                         adr2=STREET, pc=POSTCODE) %>%
        mutate(adr1=replace(adr1,which(is.na(adr1)),""))%>%
        mutate(work.adr= paste(adr1, adr2, sep=" ")) %>%
        mutate(work.adr=sub("( Romford Essex.*)| (Barking Essex.*)|( Dagenham Essex.*)|( Chadwell Heath Romford.*)|( Rush Green Romford.*)","",work.adr)) %>% #remove anything after area (e.g Barking)
        mutate(work.adr= toupper(work.adr)) %>%
        mutate(adr=paste(adr1, adr2,pc,sep=" "))# define the address to be retained

```

#### 4.2- LLPG (**ADAPTED ONLY IF LLPG is updated**).

This section do not need to be edited unless the LLPG is updated. By default, the code filters current (Approved) and historic (e.g. demolished) properties and exclude other irrelevant properties (e.g. commertial). The user could modify such filter by editting the "Logical_Status" in the code below (8=historical, 1=Approved)  . 

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}
library("dplyr")
a.working<- a %>% filter(LOGICAL_STATUS==1|8) %>% # Filtering valid addresses (e.g. non historical)
  select(UPRN=UPRN, adr= ADDRESS,pc=POSTCODE,geo_east=MAP_EAST, geo_north=MAP_NORTH,
         PAO_START_NO,PAO_END_NO) %>%
    mutate(work.adr=sub("( Romford Essex.*)|( Barking Essex.*)|( Dagenham Essex.*)|
                              ( Chadwell Heath Romford.*)|
                              ( Rush Green Romford.*)","",adr)) %>%
    mutate(work.adr= toupper(work.adr))

```

#### 5- Variables from matching dataset (b) that need to be retained. 

The names of the variables of the matching dataset that the user wants to retain after the matching process (e.g. ID, name, address) need to be specify in the code below (**use the names defined in section 4.1**).

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}

vars.b<-("id|name|surname|for.surname|dob|gen|adr")

```


##** CODE THAT DOES NOT REQUIRE MODIFICATIONS**.

## 1. Opening relevant R packages

```{r setup, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE}
library("dplyr")
library("reshape2")
library("stringr")
library("stringi")
```

## 2. Adding individual addresses for block of properties (e.g. 1-5 Mian road) in the Gazeteer (**EVALUATION==FALSE by default**)

```{r , include=FALSE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE, eval=FALSE}

# Creates sequences from the property number range
a.2<-a.working %>% filter(is.na(PAO_END_NO)==FALSE)

for (i in 1:nrow(a.2)){
  
  if (i == 1){
    se<-a.2[i,6]:a.2[i,7]
  } else {
    se2<-a.2[i,6]:a.2[i,7]
    se <- list(se,se2)
    
    no<-unlist(se)
  }        
}

# Create as many dublicate records as the length of the p.number range 
a.2 <- a.2 %>% 
  mutate(len= as.numeric(PAO_END_NO)-as.numeric(PAO_START_NO)+1)
n.times <- as.vector(a.2$len)
a.2<-a.2[rep(seq_len(nrow(a.2)), n.times),] 
a.2<- cbind(a.2,no) 

# new address replacing range by individual number
new.adr<-apply(a.2,1, function(x) {
     new.adr<-sub("[0-9]{1,3} - [0-9]{1,3}",x[10],x[8])
     new.adr<-as.vector(new.adr)
     }
  )   

#cleaning dataframe
a.2<- a.2 %>% cbind(new.adr) %>%
  mutate(work.adr= new.adr) %>%
  select(1:length(names(a.working)))


#Binding new addresses to the rest of the LLPG
a.working<- rbind(a.working,a.2)


```


## 3. Cleansing Function

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}

cleansing<- function(a.working,b.working) {

    df.list<- list(a.working,b.working)
    df <- lapply(df.list, function(x) { 
  

# Removing spaces, commas and other characteres
  
  x <- x %>% mutate(work.adr= sub("^\\s+|\\s+$", "", work.adr)) %>% #remove                               trailing and leading white space
              mutate(work.adr=  gsub("'","", work.adr)) %>% # replace apostrophe with "" 
              mutate(work.adr= gsub("[\\,\\`\\.\\_]+"," ",work.adr)) %>% # eliminates commas               and other characters
              mutate(work.adr= gsub("  "," ",work.adr)) %>% # remove double spaces (3 times               because sometimes there is more than one double space)
              mutate(work.adr= gsub("  "," ",work.adr)) %>%
              mutate(work.adr= gsub("  "," ",work.adr)) %>%
# Removing Abreviations and frequent typos
              mutate(work.adr=sub("ROD | RAOD | RD | RD$" ," ROAD ",work.adr)) %>%
              mutate(work.adr=sub(" [A-Z] ST$| [A-Z] ST ","STREET",work.adr)) %>% #change st               to street when at end of string only (#avoids the Saints!)
              mutate(work.adr=sub("ROD | RAOD | RD | RD$" ," ROAD ",work.adr)) %>%
              mutate(work.adr= sub(" AP([A-RT-XZ]+ )"," APARTMENTS ",work.adr)) %>% #correct               misspelling "appartment"          
              mutate(work.adr= sub(" AVE$| AVE "," AVENUE ",work.adr)) %>%# Sort AVENUE                   Mispelling
              mutate(work.adr= sub(" ARB([B-Z]+ )"," ARBORETUM ",work.adr)) %>%# Sort                     ARBORETUM Mispelling
              mutate(work.adr= sub(" SHRIER "," SCHRIER ",work.adr)) %>%# Sort SCHRIER                    Mispelling
              mutate(work.adr= sub(" ROPE WORKS | ROBEWORKS | ROBE WORKS "," ROPEWORKS ",
                                  work.adr)) %>%# Sort ROPEWORKS Mispelling
              mutate(work.adr= sub("ROPEWORKS ARBORETUM ","ROPEWORKS 1 ARBORETUM ",
                                  work.adr)) %>% # ADD NUMBER TO ARBORETUM PLACE
              mutate(work.adr= sub("(^FLT| FLT| FLATS|^FLATS|^FT )"," FLAT ",
                                  work.adr)) %>% # sort FLATS Mispelling
              mutate(work.adr= gsub("([0-9]+)\\s+([A-Z]\\s+)","\\1\\2",
                                   work.adr)) %>% # SORT space between flat suffix eg. 21 A change to 21A
              mutate(work.adr= gsub("0VAL","OVAL",work.adr))%>% 
# Correcting common Post Code typos
              mutate(pc=gsub(" ","",pc)) %>%
              mutate(pc= gsub("([A-Za-z]{1,2})(1)([oO])","\\110",pc))%>%
              mutate(pc= gsub("([A-Za-z])([0-9])(1)([oO])","\\1\\210",pc)) %>%
              mutate(pc= sub("^1","I",pc)) %>%
              mutate(pc= sub("^([A-Za-z]{1,2})(I)","\\11",pc))%>%
              mutate(pc= sub("^([A-Za-z]{1,2})([Oo])","\\10",pc))
  })
   a.working<<-df[[1]]
   b.working<<-df[[2]]

 }  

cleansing(a.working,b.working) 

```

## 4. Unique Identifiers Function

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}

ui<- function(a.working,b.working) {

df.list<- list(a.working,b.working)
  
df <- lapply(df.list, function(x) {

# Working Address and Post Code - (WAPC)
 x <- x %>%
  mutate(WAPC= paste0(work.adr, pc)) %>%
  mutate(WAPC= gsub(" ","",WAPC)) %>%
# Working Address Without Post Code - (WAWPC) # OJO
  mutate(WAWPC= gsub(" ","",work.adr))%>%
# Truncate first 8 characters and Post Code - (T8PC)
  mutate(T8PC= gsub("^APARTMENT|FLAT|ROOM|FLOOR","", work.adr)) %>%
  mutate(T8PC= substr(T8PC, 1, 8)) %>%
  mutate(T8PC= paste0(T8PC, pc)) %>%
  mutate(T8PC= gsub(" ","",T8PC))%>%
# Property and Premise number and Post Code - (PPNPC)
  mutate(PPNPC=(str_extract(work.adr,"[0-9]{1,3} - [0-9]{1,3}|[0-9]{1,3}-[0-9]{1,3}|[0-9]{1,3}+[a-zA-Z]{1}|[0-9]{1,3}"))) %>%
  mutate(PPNPC= gsub("-","",PPNPC))%>%
  mutate(PPNPC= ifelse(is.na(PPNPC)==TRUE,
                       paste0("NoPropNumber_",stri_rand_strings(1,4,"[a-z]")),PPNPC))%>%
  mutate(pc=ifelse(is.na(pc)==TRUE,
                   paste0("NoPostCode_",stri_rand_strings(1,4,"[a-z]")),pc))%>%
  mutate(PPNPC= paste0(PPNPC,pc)) %>% # create a unique identifier based on house
# Truncate first 10 characters- (T10)
  mutate(T10= sub("^APARTMENT|FLAT|ROOM|FLOOR","", work.adr)) %>%
  mutate(T10= gsub(" ","",T10))%>%
  mutate(T10= substr(T10, 1, 10))
 
  }
)

# Creating final dataset a (LLPG)
a.working<-df[[1]]
a.working<- a.working %>%
  select(WAPC,WAWPC,T8PC,PPNPC,T10,UPRN,adr,geo_east,geo_north)
names(a.working) <- paste0(names(a.working),"_",label.a) 

a<- a.working
fwrite(a, file = paste0(label.a,"_","tidy",".csv"))
a<<-a

# Creating final dataset b
b.working<-df[[2]]

uis<-c("WAPC|WAWPC|T8PC|PPNPC|T10|") 
vars.b<-paste0(uis,vars.b)

b.working<- b.working %>% 
  select(grep(vars.b,names(b.working))) %>%
  select(WAPC,WAWPC,T8PC,PPNPC,T10,everything())
names(b.working) <- paste0(names(b.working),"_",label.b)

b<-b.working
fwrite(b, file = paste0(label.b,"_","tidy",".csv"))
b<<-b

} 

ui(a.working,b.working)

```

## 5. Matching Function (Algorithm)

The matching algorithm is performed by running and calling the function "matching(b)", where "b" represent the dataset to be matched with the LLPG. 

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}

## Defining the variables to be retained from the LLPG
  retain<-c("UPRN_gz","adr_gz","geo_east_gz","geo_north_gz")
## Steps order
  steps<-c(1,2,3,5,4)    
 
matching<-function(x) { 

  a<- a %>% mutate(No=1:nrow(a))
  b<- x %>% mutate(No=c(1:nrow(x)),method=NA)

# First Matching    
join<-merge(b,a,by.y=names(a)[steps[1]], 
            by.x=names(b)[steps[1]], all.x=TRUE) 

join <- join[!duplicated(join$No.x),]# remove duplicates from first join

join$method[join$No.y!="NA"]<-names(a)[steps[1]]# Updates matching method

join<- join %>% select(1:length(names(b)),method,No=No.y,-No.x) %>%
      mutate(index=seq(1:nrow(join)))

total<- join 

  for (i in 2:length(steps)){ # loop to match based on each unique identifier

filt<-  total %>% filter(is.na(method)==TRUE) %>%
            select(index) %>% distinct() %>% unlist() 

data<- total[filt,] 

join<-merge(data,a,by.y=names(a)[steps[i]], 
                by.x=names(b)[steps[i]], all.x=TRUE) %>%
      mutate(valid=ifelse(is.na(No.y)==FALSE,"valid",NA)) %>%
      select(valid,index) %>% 
      unique()

total<- left_join(total,join,by="index") 

total$method[total$valid=="valid"]<-names(a)[steps[i]] 
total<- total %>% select(-valid) 

  }     

final<-left_join(total,a,by="No") %>% select(names(b),retain)

##################################
# address-specific direct match ##
##################################

  adr.b<- names(b)[grep("work.adr",names(b))] 

matched<- final %>% 
    mutate(UPRN_def=ifelse(grepl("ALEX[A_Z]+.+ (COURT|CARE|320)|320 RAIN",
      get(adr.b))==TRUE,  "010023591471",
                ifelse(grepl("ABBEY CARE HOME",
      get(adr.b))==TRUE,"000100003891",
                ifelse(grepl("PARKVIEW [NC]|PARK VIEW N|1-2 MORL|PARK VIEW CARE HOME", get(adr.b))==TRUE,"000100045662",
                ifelse(grepl("CHASEVIEW|CHASE VIEW", get(adr.b))==TRUE,"000100101940",
                ifelse(grepl("YMCA", get(adr.b))==TRUE,"000100102744",UPRN_gz)))))) %>%
    mutate(method=ifelse(is.na(method)==TRUE &
                            is.na(UPRN_def)==FALSE,"direct.match",
                         ifelse(is.na(method)==TRUE &
                            is.na(UPRN_def)==TRUE,"no.match",method)))
##########
# OUTPUT #
##########
fwrite(matched, file = "matched.csv")
matched<<-matched
summary<-matched %>% group_by(method) %>% count() %>% arrange(desc(n))

print(as.data.frame(summary)) 

# Message
msg.mch<- nrow(b)- summary[grep("no.match",summary$method),2]
msg.tot<-ifelse(nrow(b)==sum(summary$n),nrow(b),"ERROR!")  

effectiveness.index<<-round(msg.mch/msg.tot*100,2) 
msg<-paste0("The algorithm matched ", msg.mch, " out of ",msg.tot, " records. This represents a matching Effectiveness of ",effectiveness.index,"%.")   

print(msg)

}     

matching(b)

```

