---
title: "matching_by_names_pilot"
author: "Juan Carpio"
date: "21 December 2017"
output: html_document
---

```{r setup, include=FALSE}
require("knitr")
opts_knit$set(root.dir = "//dscs/cscom/Policy/Corp Information/Olypmic Hosts NKM (secure)/Community Mapping 2017/Juans_tests/Juan_2017")

#knitr::opts_chunk$set(echo = TRUE)
```



# Define working Directory
```{r}
setwd("//dscs/cscom/Policy/Corp Information/Olypmic Hosts NKM (secure)/Community Mapping 2017/Juans_tests/Juan_2017")

```


## Opening relevant R packages

```{r}
library("data.table")
library("dplyr")
library("reshape2")
library("stringi")

rm(list = ls())

```

## Uploading datasets
```{r}
gp<- fread("gp_matched.csv",colClasses=(rep(c("character"),22))) # GP data
sch<- fread("sch_matched.csv",colClasses=(rep(c("character"),20))) #school data

postcodes<- fread("//dscs/cscom/Policy/Corp Information/Olypmic Hosts NKM (secure)/Community Mapping 2017/Juans_tests/2015/postcodes.csv") # Post codes within the Borough bounderies

```

### 3.2- Example filtering records with Post Codes out of the Borough (eval=FALSE by default)

Filtering addresses out of the Borough is crucial for some data sets such as the School Census.

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE, eval=FALSE}

sch <- sch %>% 
  mutate(pc= str_extract(adr_sch,"^([Gg][Ii][Rr] 0[Aa]{2})|((([A-Za-z][0-9]{1,2})|(([A-Za-z][A-Ha-hJ-Yj-y][0-9]{1,2})|(([A-Za-z][0-9][A-Za-z])|([A-Za-z][A-Ha-hJ-Yj-y][0-9]?[A-Za-z])))) [0-9][A-Za-z]{2})$"))%>% # UK regex to validate postcodes
  mutate(pc=gsub(" ","",pc))

colnames(postcodes)[2] <- "pc"

nrow(sch)#41991
sch<- semi_join(sch,postcodes, by="pc") 
nrow(sch)#

```



## Define variable labels

```{r , include=TRUE, warning=FALSE, cache=TRUE, message=FALSE,echo=FALSE}

label.a<- "gp" # GP
  
label.b<- "sch" # School
  
```


## Selecting only relevant variables
```{r}
criteria<-c("id|dob|name|surname|gen|UPRN.def")

gp_tidy<-select(gp,matches(criteria))
sch_tidy<-select(sch,matches(criteria))


```


## Cleansing and UIs creation
```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}


df.list<- list(gp_tidy,sch_tidy)
 
cleansing <- lapply(df.list, function(x) {  
  
  names(x)<- gsub("_.*","",names(x))
  
  x<- x %>%
    #Cleaning
    mutate(UPRN=ifelse(UPRN=="",paste0("NA_",stri_rand_strings(1,4,"[a-z]")),
                       UPRN)) %>%
    mutate(name=ifelse(name=="",paste0("NA_",stri_rand_strings(1,4,"[a-z]")),
             toupper(name))) %>%
    mutate(name=gsub("[\\,\\`\\'\\.]","",name)) %>% #remove random characters 
    mutate(surname=ifelse(surname=="",
              paste0("NA_",stri_rand_strings(1,4,"[a-z]")),
              toupper(surname))) %>%
    mutate(surname=gsub("[\\,\\`\\'\\.]","",surname)) %>% 
    mutate(for.surname=ifelse(sum(grepl("for.surname",names(gp_tidy)))>0 &
                  for.surname!="", toupper(for.surname),
                  paste0("NA_",stri_rand_strings(1,4,"[a-z]")))) %>%    
    mutate(for.surname=ifelse(sum(grepl("for.surname",names(gp_tidy)))>0,
                  toupper(for.surname),
                  paste0("NA_",stri_rand_strings(1,4,"[a-z]")))) %>%    
    # UNSD - UPRN, Name,Surname, dob
    mutate(UNSD=paste0(UPRN,name,surname,dob)) %>%
    # UNFSD - UPRN, Name,Former Surname, dob  
    mutate(UNFSD=paste0(UPRN,name,for.surname,dob)) %>%
    # UISND - UPRN, Inverted Surname - Name, dob  
    mutate(UISND=paste0(UPRN,surname,name,dob)) %>%
    # UIFSND -  UPRN, Inverted Former Surname - Name, dob  
    mutate(UIFSND=paste0(UPRN,for.surname,name,dob)) %>%
    # USNSD - UPRN, Single Name, Single Surname, dob  
    mutate(USNSD=paste0(UPRN,gsub("-.*| .*","",name),
                        gsub("-.*| .*","",surname),dob)) %>%
    # USNFSSD - UPRN, Single Name, Single Former Surname, dob  
    mutate(USNFSSD=paste0(UPRN,gsub("-.*| .*","",name),
                        gsub("-.*| .*","",for.surname),dob)) %>%
    # UISSSND - UPRN, Inverted Single Surname - Single Name,dob  
    mutate(UISSSND=paste0(UPRN, gsub("-.*| .*","",surname),
                          gsub("-.*| .*","",name), dob)) %>%
    # UISFSSND - UPRN, Inverted Single FORMER Surname - Single Name,dob  
    mutate(UISFSSND=paste0(UPRN, gsub("-.*| .*","",for.surname),
                          gsub("-.*| .*","",name), dob)) %>%
    # BIUNSD - UPRN, Bigram Name, Bigram Surname, dob
    mutate(BIUNSD=paste0(UPRN,substr(name,1,2),substr(surname,1,2),dob)) %>%
    # BIUNFSD - UPRN, Bigram Name, Bigram Former Surname, dob  
    mutate(BIUNFSD=paste0(UPRN,substr(name,1,2),
                        substr(for.surname,1,2),dob)) %>%
    # UND - UPRN, Name, dob
    mutate(UND=paste0(UPRN,name,dob)) %>%
    # USD - UPRN, Name, dob
    mutate(USD=paste0(UPRN,surname,dob)) %>%
    # UFSD - UPRN, Former Surname, dob  
    mutate(UFSD=paste0(UPRN,for.surname,dob)) %>%
    #### NO UPRN ####
    # GNSD - Gender, Name,Surname, dob
    mutate(GNSD=paste0(gen,name,surname,dob)) %>%
    # GNFSD - Gender Name,Former Surname, dob  
    mutate(GNFSD=paste0(gen,name,for.surname,dob)) %>%
    # GISND - Gender, Inverted Surname - Name, dob  
    mutate(GISND=paste0(gen,surname,name,dob)) %>%
    # GIFSND -  Gender, Inverted Former Surname - Name, dob  
    mutate(GIFSND=paste0(gen,for.surname,name,dob)) %>%
    # GTRNSD - Gender, Trigram Name, Trigram Surname, dob
    mutate(GTRNSD=paste0(gen,substr(name,1,3),substr(surname,1,3),dob))%>%
    # GTRNFSD - Gender, Trigram Name, Trigram Surname, dob
    mutate(GTRNFSD=paste0(gen,substr(name,1,3),
                          substr(for.surname,1,3),dob))%>%
    select(8:26,1:7)# this is to put the UIs at the beginning                                    # of the data.frame

   
    } ) 

    a<-as.data.frame(cleansing[[1]])
    names(a) <- paste0(names(a),"_",label.a)
    
    b<-as.data.frame(cleansing[[2]])
    names(b) <- paste0(names(b),"_",label.b)

```


## 5. Matching Function (Algorithm)

The matching algorithm is performed by running and calling the function "matching(b)", where "b" represent the dataset to be matched with the LLPG. 

```{r , include=TRUE, warning=FALSE, cache=FALSE, message=FALSE,echo=FALSE}

## Defining the variables to be retained
  retain<-c("id","name","surnmae","dob")

  steps_a<-c(1:19)    
  steps_b<-c(1,2,1,2,5,6,5,6,9,10,11,12,13,14,15,14,15,18,19)    

matching<-function(x,y) { 

  a<- x %>% mutate(No=1:nrow(x))
  b<- y %>% mutate(No=c(1:nrow(y)),method=NA)


# First Matching    
join<-merge(b,a,by.y=names(a)[steps_a[1]], 
            by.x=names(b)[steps_b[1]], all.x=TRUE) 

join <- join[!duplicated(join$No.x),]# remove duplicates from first join

join$method[join$No.y!="NA"]<-names(a)[steps_a[1]]# Updates matching method

join<- join %>% select(1:length(names(b)),method,No=No.y,-No.x) %>%
      mutate(index=seq(1:nrow(join)))

total<- join 
#nuevo test
#names(total)<-gsub("\\..*","",names(total))

  for (i in 2:length(steps_a)){ #loop to match based on each unique identifier

filt<-  total %>% filter(is.na(method)==TRUE) %>%
            select(index) %>% distinct() %>% unlist() 

data<- total[filt,] 

join<-merge(data,a,by.y=names(a)[steps_a[i]],#cambio i 
                by.x=names(b)[steps_b[i]], all.x=TRUE) %>% #cambio i
      mutate(valid=ifelse(is.na(No.y)==FALSE,"valid",NA)) %>%
      select(valid,index) %>% 
      unique()

total<- left_join(total,join,by="index") 

total$method[total$valid=="valid"]<-names(a)[steps_a[i]] #cambio i
total<- total %>% select(-valid) 

  }     

final<-left_join(total,a,by="No") 
final<-final %>% select(grep("id|name|dob|method",names(final))) %>%
      mutate(method=ifelse(is.na(method)==TRUE,"no.match",method))
  #select(names(b),retain)


##########
# OUTPUT #
##########
fwrite(final, file = "matched.csv")
matched<<-final
summary<-final %>% group_by(method) %>% count() %>% arrange(desc(n))

print(as.data.frame(summary)) 

# Message
msg.mch<- nrow(b)- summary[grep("no.match",summary$method),2]
msg.tot<-ifelse(nrow(b)==sum(summary$n),nrow(b),"ERROR!")  

effectiveness.index<<-round(msg.mch/msg.tot*100,2) 
msg<-paste0("The algorithm matched ", msg.mch, " out of ",msg.tot, " records. This represents a matching Effectiveness of ",effectiveness.index,"%.")   

print(msg)

}     

matching(a,b)

```
